Properties of Methods for Assessing Replication

### 2018-04-18 Notes on McNemar's Test.Rmd

## A framework for replication
Let there be $k$ pairs of studies each containing an original study and a replication study. Each pair of studies can be characterized by a vector of true effect parameters $(\theta_{1,j}, \theta_{2,j})$ that are unobserved with zero sampling variance. Each pair of studies is further associated with estimates $(T_{1,j}, T_{2,j})$ of these true effects. These estimates have sampling variances $V_{1,j}$ and $V_{2,j}$ which are assumed to be known. Let $T_{i,j} \sim N(\theta_{i,j}, V_{i,h})$ for $i=1,2$ and $j=1,2,\ldots, k$. Further, assume that $T_{i,j}$ are independent. Generally speaking, two studies replicate if $\theta_{1,j} \approx \theta_{2,j}$.\footnote{We observe \textit{strict} replication when $\theta_{1,j} = \theta_{2,j}$.}

@@

The Open Science Foundation (OSF) propose four different test for assessing replication based on significance and p-values. These methods include Fish'er method, McNemar's test, the Wilcoxon signed-rank test, and the \textit{t} test for dependent samples. In deriving properties for each of these tests, we assume a two-sided hypothesis test for each of the $2k$ studies with a significance level of 0.05. Let $p_{1,j}$ denote the p-value of the i\textsuperscript{th} original study and let $p_{2,j}$ denote the p-value of the i\textsuperscript{th} replication study.

### The Umbrella of the Sign Test

Before diving into the technical details, it may be worth making a quick observation about the relationship between three of these tests. As shown in the diagram below, three of these tests can be thought of as an extension on the sign test. If we chose to test for differences in pairs of p-values using the sign test, we would let $\rho = P(p_{1,i} > p_{2,i})$ and test $H_0: \rho = 0.5$. In this case, we don't make use of the actual p-values but rather the \textit{sign} of the difference in p-values. Suppose we want to use a little more information. We might observe that p-values can be ranked, which would call for use of the Wilcoxon signed-rank test. Alternatively, we might observe that we can categorize p-values as being either significant or non-significant. In this case, our observations are binary and we can use McNemar's test. Finally, we might make full use of the "information" contained in the p-values, acknowledging that they are continuous variables and assuming that the difference in p-values is approximately normally distributed. In this case, we might gravitate towards the t test.

This framework for the relationship for these tests may, or may not, prove to be useful in understanding the properties of these tests in the context of replication. It could be that this relationship can help explain congruence between the tests. It could also be the case that since these tests are using different amounts of information about the p-values, that they may produce results that do not necessarily confirm one another.

 ```{r, echo=FALSE}
par(mar=c(1,1,1,1), mfrow=c(1,1))

names <- c("Sign Test", "Wilcoxon Test", "t Test", "McNemar's Test")
M <- matrix(nrow=4, ncol=3, byrow=T, data=0)

M[2,1] <- "Ranked"
M[3,1] <- "Normal"
M[4,1] <- "Binary"
plotmat(M, pos = c(1, 3), curve=0, name = names, lwd = 1,
        box.lwd = 2, cex.txt = 0.6, dtext=-0.9, box.size = 0.15,
        box.type = "square", box.prop = 0.5)

```

@@

[Something to note, but for right now, I won't be making use of it...]

Back to $\beta_M$: If we assume that the original and replication studies are independent of one another, then $p_{10} = \frac{1}{k} \sum\limits_{j=1}^k (1-\beta_{1j})\beta_{2j} $ and $p_{01} = \frac{1}{k}\sum\limits_{j=1}^k \beta_{1j}(1-\beta_{2j})$ where $\beta_{1j}$ is the power of the j\textsuperscript{th} original study and $\beta_{2j}$ is the power of the j\textsuperscript{th} replication study.

One thing we may want to consider is what happens as the replication sample sizes grow large. If $n_{2j} \rightarrow \infty$ $\forall j$, then $\beta_{2j} \rightarrow 0$ (or power goes to 1) implying that $p_{10} \rightarrow \frac{1}{k} \sum\limits_{j=1}^k \beta_{1j}$ and $p$

>>>>

## A framework for replication
Let there be $k$ pairs of studies each containing an original study and a replication study. Each pair of studies can be characterized by a vector of true effect parameters $(\theta_{1,i}, \theta_{2,i})$ that are unobserved with zero sampling variance. Each pair of studies is further associated with estimates $(T_{1,i}, T_{2,i})$ of these true effects. These estimates have sampling variances $V_{1,i}$ and $V_{2,i}$ which are assumed to be known. Let $T_{j,i} \sim N(\theta_{j,i}, V_{j,i})$ for $j=1,2$ and $i=1,2,\ldots, k$. Further, assume that $T_{j,i}$ are independent. Generally speaking, two studies replicate if $\theta_{1,i} \approx \theta_{2,i}$.

@@

The authors propose four different tests for assessing replication based on significance and p-values. These methods include Fisher's method, McNemar's test, the Wilcoxon signed-rank test, and the \textit{t} test for dependent samples. In what follows, we derive properties of each of these tests and discuss how they function within our proposed framework for replication.

The authors assume a two-sided hypothesis test for each of the $2k$ studies with a significance level of 0.05. Let $p_{1,i}$ denote the p-value of the i\textsuperscript{th} original study and let $p_{2,i}$ denote the p-value of the i\textsuperscript{th} replication study.

@@
